Restrictive Clearance Comparison

### 2024-11-05 data-raw/RestrictiveClearance_comparison.Rmd
Adding new data-raw/ files prior to merge

```{r library_loading}
devtools::load_all()
library(httk)
library(tidyverse)
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
httk::load_dawson2021()
```

@@

```{r dawson}
# try dawson
# can also be run as background job (see data-raw/RC_comparison_bkgd.R)

human_chems_gas <- get_cheminfo(info = "DTXSID",
                                species = "Human",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)

common_chems_dawson <- cvt %>%
  filter(species %in% c("human", "rat", "mouse"),
         analyte_dtxsid %in% human_chems_gas) %>%
  pull(analyte_dtxsid) %>%
  unique()

rat_chems_gas <- get_cheminfo(info = "DTXSID",
                                species = "Rat",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)

common_chems_rat <- cvt %>%
  filter(species %in% "rat",
         analyte_dtxsid %in% rat_chems_gas) %>%
  pull(analyte_dtxsid) %>%
  unique()

remove(list = c("human_chems_gas", "rat_chems_gas"))

# How many are in the previous invivoPKfit analysis?
length(common_chems_dawson %in% unique(my_pk$fit %>%
                                         filter(convcode == 0) %>%
                                         pull(Chemical)))

test_parameterization <- function(dtxsid, species = "Human") {
  is(
    tryCatch(
      {
        parameterize_gas_pbtk(dtxsid = dtxsid,
                              species = species,
                              supress.messages = TRUE)  
      }, error = function(e) {
        NULL
      }
    ),
    "list"
  )
}

parameterize_all_gas <- function(dtxsid, species = "Human") {
  # Note that Clint is given in uL/min, which must be converted to L/h
  params <- httk::parameterize_gas_pbtk(dtxsid = dtxsid,
                                        species = species,
                                        default.to.human = FALSE)
  paramdf <- data.frame(Chemical = dtxsid, Species = tolower(species),
                        Qliverf = params$Qliverf,
                        Funbound.plasma = params$Funbound.plasma,
                        Clint = params$Clint*60/1e+6)
  paramdf$fupClint <- paramdf$Funbound.plasma * paramdf$Clint
  paramdf$ExtEff <- paramdf$fupClint/(paramdf$Qliverf + paramdf$fupClint)
  paramdf$Q_Clint <- paramdf$Qliverf/paramdf$Clint
  return(paramdf)
}

if (all(vapply(common_chems_rat, test_parameterization,
               FUN.VALUE = logical(1), species = "Rat"))) {
  rat_params <- do.call(rbind, lapply(common_chems_rat,
                                      parameterize_all_gas,
                                      species = "Rat"))
}

if (all(vapply(common_chems_dawson, test_parameterization,
               FUN.VALUE = logical(1), species = "Human"))) {
  human_params <- do.call(rbind, lapply(common_chems_dawson,
                                        parameterize_all_gas,
                                        species = "Human"))
}
```

@@

```{r get_clearances}
params_dtxsid <- function(dtxsid, species) {
  httk::parameterize_gas_pbtk(dtxsid = dtxsid,
                              species = species,
                              default.to.human = FALSE,
                              suppress.messages = TRUE)
}

make_clearance_df <- function(dtxsid, species) {
  tmp_df <- data.frame(Chemical = dtxsid)
  tmp_params <- lapply(dtxsid, params_dtxsid, species = species)
  tmp_df$restrict_cl <- vapply(tmp_params,
                               \(x) {
                                 calc_total_clearance(
                                   parameters = x,species = species,
                                   suppress.messages = TRUE,
                                   restrictive.clearance = TRUE
                                 )
                               },
                               FUN.VALUE = double(1L)
  )
  tmp_df$nonrestrict_cl <- vapply(tmp_params,
                                  \(x) {
                                    calc_total_clearance(
                                      parameters = x,species = species,
                                      suppress.messages = TRUE,
                                      restrictive.clearance = FALSE
                                    )
                                  },
                                  FUN.VALUE = double(1L)
  )
  return(tmp_df)
}
clearance_df <- make_clearance_df(dtxsid = common_chems_rat, species = "rat")
```

>>>>

```{r library_loading}
devtools::load_all()
library(httk)
library(tidyverse)
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
httk::load_dawson2021()
source("data-raw/RC_comparison_bkgd.R")
```

@@

@@
